Reliable error estimation for Sobol' indices

### 2016-03-10 replicated_sobol.tex
Adding more on Sobol background and definition of replicated sequences

\newcommand{\fudge}{\fC}
\newcommand{\dtf}{\textit{\doubletilde{f}}}
\newcommand{\cube}{[0,1)^d}
%\renewcommand{\bbK}{\natzero^d}
\newcommand{\rf}{\mathring{f}}
\newcommand{\rnu}{\mathring{\nu}}
\newcommand{\natm}{\naturals_{0,m}}
\newcommand{\wcS}{\widecheck{S}}
\newcommand{\tol}{\text{tol}}
\newcommand{\e}{\text{e}}
\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\vx}{\bvec{x}}
\newcommand{\vi}{\bvec{i}}
\newcommand{\ve}{\bvec{e}}
\newcommand{\vk}{\bvec{k}}
\newcommand{\vz}{\bvec{z}}
\newcommand{\dif}{\mathsf{d}}
\newcommand{\hf}{\hat{f}}
\newcommand{\hS}{\widehat{S}}
\newcommand{\tS}{\widetilde{S}}
\newcommand{\tf}{\tilde{f}}
\newcommand{\fC}{\mathfrak{C}}
\newcommand{\homega}{\widehat{\omega}}
\newcommand{\wcomega}{\mathring{\omega}}
\newcommand{\vzero}{\bvec{0}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}

@@

\section{Problem Statement}
Denote by $\vx \in [0,1)^d$ a point with components $x_1,\dots,x_d$. Let $u$ be a subset of $\{1,\dots,d\}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^u$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, we define the following hybrid point: 
\[(\vx_u:{\vx'}_{-u}) := ({x'}_1,\dots,{x'}_{u-1},x_u,{x'}_{u+1},\dots,{x'}_d).\]
Let $f$ be the symbol representing the numerical model considered. We assume $f \in \mathcal{L}^2[0,1]^d$. Denote by $\mu$ and $\sigma^2$ the mean and variance of $f$.
%These two quantities can be expressed as the following integrals:
%\[\mu= \frac{1}{2} \int_{[0,1)^{2d-1}}f(\vx)+f(\vx_u:{\vx'}_{-u}) d\vx d{\vx'}_{-u}, \]
%\[\sigma^2= \int_{[0,1)^{d}}f(\vx)^2 d\vx - \Big{(}\int_{[0,1)^{d}}f(\vx) d\vx\Big{)}^2 .\]
The uncertainty on $\vx$ is modeled by random variables such that $\vx \sim \mathcal{U}[0,1]^d$. Recall the Hoeffding decomposition of $f$:
\begin{equation}
f(\vx)=\mu+\sum \limits_{u \subseteq \{1,\dots,d\}} f_u(\vx),
\label{anova}
\end{equation}
where:
\[f_u(\vx)= \int_{[0,1)^{|u|}} f(\vx) d{\vx}_{-u} - \sum \limits_{v \subset u} f_v(\vx).\]
Due to orthogonality, by taking the variance of each side in Equation \ref{anova} we obtain the following decomposition of the variance:
\[ \sigma^2 = \sum \limits_{u \subseteq \{1,\dots,d\}} \sigma_v^2.\]
The problem of interest is the estimation of the quantities:
\[\underline{\tau}_u^2 = \sum \limits_{v \subseteq u} \sigma_v^2, \qquad \text{ for } u = 1,\dots,d.\]
These quantities correspond to the partial sensitivity Sobol' indices often found written in the literature in their normalized form: $\underline{\tau}_u^2/\sigma^2$. When, $|u|=1$, $\underline{\tau}_u^2$ corresponds to the first-order Sobol' index that evaluate the main effect of the variable $\vx_u$. In \cite{}, Owen proposed to use the following expression for $\underline{\tau}_u^2$:
\[\underline{\tau}_u^2  =\int_{[0,1)^{2d-1}}(f(\vx) - \mu) (f(\vx_u:{\vx'}_{-u})-\mu) d\vx d{\vx'}_{-u},\]
taking $\mu$ as introduced in Janon \textit{et al.}:
\[\mu= \frac{1}{2} \int_{[0,1)^{2d-1}}f(\vx)+f(\vx_u:{\vx'}_{-u}) d\vx d{\vx'}_{-u}. \]
The natural way to estimate these two quantities is for $(\vx_i,\vx'_i) \stackrel{iid}{\sim} [0,1)^{2d}$ via:
\begin{align*}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=1}^n (f(\vx_i) - \mu) (f(\vx_{i,u}:{\vx'}_{i,-u})-\mu),\\
\widehat{\mu} & = \frac{1}{2n} \sum \limits_{i=1}^n f(\vx_i) +f(\vx_{i,u}:{\vx'}_{i,-u}).
\end{align*}
\bigskip

@@

$\forall \ u \subsetneq \{1,\dots,d\}$ such that $|u|=r$, $X_u$ and $X'_u$ represent the same point sets.
\end{defin}
The replication procedure introduced in \cite{} allows to estimate all first-order Sobol' indices with two replicated designs of order $1$. In this procedure, the quantity $f(\vx_u:{\vx'}_{-u})$ is only evaluated $n$ times once for all $u$ resulting in a total of $2n$ evaluations of $f$. Denote by $X$,$X'$ the two replicated designs of order $1$ required by the replication procedure, $X'$ is constructed from $X$ as follows:
\[{x'}_{i,j}=x_{\pi_j(i),j}, \qquad \forall \ (i,j) \in \{1,\dots,n\} \times \{1,\dots,d\},\]
where $\pi_1,\dots,\pi_d$ are $d$ permutations on $\{1,\dots,n\}$.
\bigskip

@@

\section{Sobol' sequence background}
Obviously, we do not need to evaluate $f(\vx)$ $d$ times. But we can do event better than that, and evaluate $f(\vx_u:{\vx'}_{-u})$ only one time. This is the replication method....

@@

\section{Matrix generators group $U_m$}

@@

>>>>

\newcommand{\fudge}{\fC}
\newcommand{\dtf}{\textit{\doubletilde{f}}}
\newcommand{\cube}{[0,1)^d}
\newcommand{\cubes}{[0,1)^s}
%\renewcommand{\bbK}{\natzero^d}
\newcommand{\rf}{\mathring{f}}
\newcommand{\rnu}{\mathring{\nu}}
\newcommand{\natm}{\naturals_{0,m}}
\newcommand{\wcS}{\widecheck{S}}
\newcommand{\tol}{\text{tol}}
\newcommand{\e}{\text{e}}
\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\vx}{\bvec{x}}
\newcommand{\vi}{\bvec{i}}
\newcommand{\ve}{\bvec{e}}
\newcommand{\vk}{\bvec{k}}
\newcommand{\vz}{\bvec{z}}
\newcommand{\dif}{\mathsf{d}}
\newcommand{\hf}{\hat{f}}
\newcommand{\hS}{\widehat{S}}
\newcommand{\tS}{\widetilde{S}}
\newcommand{\tf}{\tilde{f}}
\newcommand{\fC}{\mathfrak{C}}
\newcommand{\homega}{\widehat{\omega}}
\newcommand{\wcomega}{\mathring{\omega}}
\newcommand{\vzero}{\bvec{0}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}

@@

\section{Problem Statement}
Denote by $\vx \in [0,1)^d$ a point with components $x_1,\dots,x_d$. Let $u$ be a subset of $\{1,\dots,d\}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^u$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, we define the following hybrid point: 
\[(\vx_u:{\vx'}_{-u}) := ({x'}_1,\dots,{x'}_{u-1},x_u,{x'}_{u+1},\dots,{x'}_d).\]
Let $f$ be the symbol representing the numerical model considered. We assume $f \in \mathcal{L}^2[0,1]^d$. Denote by $\mu$ and $\sigma^2$ the mean and variance of $f$.
%These two quantities can be expressed as the following integrals:
%\[\mu= \frac{1}{2} \int_{[0,1)^{2d-1}}f(\vx)+f(\vx_u:{\vx'}_{-u}) d\vx d{\vx'}_{-u}, \]
%\[\sigma^2= \int_{[0,1)^{d}}f(\vx)^2 d\vx - \Big{(}\int_{[0,1)^{d}}f(\vx) d\vx\Big{)}^2 .\]
The uncertainty on $\vx$ is modeled by random variables such that $\vx \sim \mathcal{U}[0,1]^d$. Recall the Hoeffding decomposition of $f$:
\begin{equation}
f(\vx)=\mu+\sum \limits_{u \subseteq \{1,\dots,d\}} f_u(\vx),
\label{anova}
\end{equation}
where:
\[f_u(\vx)= \int_{[0,1)^{|u|}} f(\vx) d{\vx}_{-u} - \sum \limits_{v \subset u} f_v(\vx).\]
Due to orthogonality, by taking the variance of each side in Equation \ref{anova} we obtain the following decomposition of the variance:
\[ \sigma^2 = \sum \limits_{u \subseteq \{1,\dots,d\}} \sigma_v^2.\]
The problem of interest is the estimation of the quantities:
\[\underline{\tau}_u^2 = \sum \limits_{v \subseteq u} \sigma_v^2, \qquad \text{ for } u = 1,\dots,d.\]
These quantities correspond to the partial sensitivity Sobol' indices often found written in the literature in their normalized form: $\underline{\tau}_u^2/\sigma^2$. When, $|u|=1$, $\underline{\tau}_u^2$ corresponds to the first-order Sobol' index that evaluate the main effect of the variable $\vx_u$. In \cite{}, Owen proposed to use the following expression for $\underline{\tau}_u^2$:
\begin{equation}\label{first_order_indice}
\underline{\tau}_u^2  =\int_{[0,1)^{2d-1}}(f(\vx) - \mu) (f(\vx_u:{\vx'}_{-u})-\mu) d\vx d{\vx'}_{-u},
\end{equation}
taking $\mu$ as introduced in Janon \textit{et al.}:
\[\mu= \frac{1}{2} \int_{[0,1)^{2d-1}}f(\vx)+f(\vx_u:{\vx'}_{-u}) d\vx d{\vx'}_{-u}. \]
The natural way to estimate these two quantities is for $(\vx_i,\vx'_i) \stackrel{iid}{\sim} [0,1)^{2d}$ via:
\begin{align}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=1}^n (f(\vx_i) - \widehat{\mu}) (f(\vx_{i,u}:{\vx'}_{i,-u})-\widehat{\mu}),\label{first_order_indice_estimator}\\
\nonumber
\widehat{\mu} & = \frac{1}{2n} \sum \limits_{i=1}^n f(\vx_i) +f(\vx_{i,u}:{\vx'}_{i,-u}).
\end{align}
\bigskip

@@

$\forall \ u \subsetneq \{1,\dots,d\}$ such that $|u|=r$, $X_u$ and $X'_u$ represent the same point sets.
\end{defin}
The replication procedure introduced in \cite{•} allows to estimate all first-order Sobol' indices with two replicated designs of order $1$. In this procedure, the quantity $f(\vx_u:{\vx'}_{-u})$ is only evaluated $n$ times once for all $u$ resulting in a total of $2n$ evaluations of $f$. Denote by $X$,$X'$ the two replicated designs of order $1$ required by the replication procedure, $X'$ is constructed from $X$ as follows:
\[{x'}_{i,j}=x_{\pi_j(i),j}, \qquad \forall \ (i,j) \in \{1,\dots,n\} \times \{1,\dots,d\},\]
where $\pi_1,\dots,\pi_d$ are $d$ permutations on $\{1,\dots,n\}$.
\bigskip

@@

\section{Digital Nets Background}

In this section we introduce a particular type of quasi-Monte Carlo points to estimate $\eqref{first_order_indice}$, which will allow us to evaluate the integrand at points $\{\left(\vx_i,{\vx'}_{i,-u}\right)\}_{i=0}^{2^m}$ only one time to estimate all $d$ indices.

The concept of $(t,m,s)$-net and $(t,s)$-sequence was first introduced and studied by Niederreiter in \cite{•}. These are points sets in $\cubes$ whose quality is measured by the parameter $t$. Because this parameter is the criterion we use to construct our new sequences, we provide the key definitions below.

Let $\mathcal{A}$ be the set of all $A\in\cubes$ such that $A:=\prod_{i=1}^s [\alpha_jb^{-\gamma_j},(\alpha_j+1)b^{-\gamma_j})$, with integers $0\leq\gamma_j$ and $0\leq\alpha_j<b^{\gamma_j}$ for all $j=1,\dots,s$.

\begin{defin}
For integers $s\geq 1$, $b\geq 2$, and $m\geq t\geq 0$, the set of points $\mathcal{P}\in\cubes$, with $|\mathcal{P}|=b^m$, is a $(t,m,d)-net$ in base $b$ if for every $A\in\mathcal{A}$ with volume $b^{t-m}$, $|A\cap\mathcal{P}|=b^t$ ($A$ contains exactly $b^t$ points of $\mathcal{P}$).
\end{defin}

In other words, any box of ... contains a proport
....
This definition is linked to the star discrepancy...
The smaller the t value, the better.

\begin{defin}
For integers $s\geq 1$, $b\geq 2$, and $t\geq 0$, the sequence $\{\vx_i\}_{\mathbb{N}_0}$ is a $(t,s)$-sequence in base $b$, if every set $\mathcal{P}_{l,m}:=\{\vx_i\}_{i=lb^m}^{(l+1)b^m-1}$ with $l\geq 0$ and $m\geq t$, $\mathcal{P}_{l,m}$ is a $(t,m,s)$-net in base $b$.
\end{defin}

Digital nets are usually computed with the generating matrices $C_1,\dots,C_s$ in $\mathbb{F}_b$, whose size $\infty\times\infty$. However, to construct the first $b^m$ points, one only needs the knowledge of $C_1^{\infty\times m},\dots,C_s^{\infty\times m}$. For each $i=0,\dots,b^m-1$, the points $\vx_i = (x_{i,j},\dots,x_{i,s})^\intercal$ of the sequence can be constructed as follows:
\begin{equation}
(x_{i,j,1},x_{i,j,2},\dots)^\intercal = C_j^{\infty\times m}(i_{0},\dots,i_{m-1})^\intercal,\qquad j= 1,\dots,s\, ,
\end{equation}
where $x_{i,j} = \sum_{k \geq 1}x_{i,j,k}b^{-k}$ and $i = \sum_{k=0}^{m-1}i_kb^{k}$.

Our goal is to generate two sequences $\{\vx_i\}_{i\in\mathbb{N}_0}$ and $\{{\vx'}_i\}_{i\in\mathbb{N}_0}$ such that for all $m\geq 0$, $\{{\vx'}_i\}_{i=0}^{b^m-1}$ is a replicated design of $\{{\vx}_i\}_{i=0}^{b^m-1}$. For that purpose, $C_j'$ depend on the choice of $C_j$.

Although the choices of $C_j$ and $C_j'$ are many, there is a special case interesting for its construction simplicity, and with a wide range of possible $C_j'$ choices given $C_j$.

Considering the generating matrices to be upper triangular full rank (no zeros on the diagonal), one ensures that $x_{i,j,k}=0$, for $k>m$ and $b^m-1\geq i \geq 0$. Since the block $C_j^{m\times m}$ is full rank, each $x_{i,j}\neq x_{r,j}$ when $i\neq j$. Therefore, excluding the case ${\vx}_i={\vx'}_i$ for all $i\in\mathbb{N}_0$, it is enough to choose $C_j\neq C_j'$ for at least one dimension $j$. When $b=2$, this is the case of Sobol' sequences and one could simply consider $C_{j+d},\dots,C_{2d}$ to generate the replicated sequence.

Because we are estimating the integral \eqref{first_order_indice}, we want to minimize the t-value of the following $d$ sequences at the same time............

According to \eqref{first_order_indice}, our dimension of study is $s = 2d-1$.

@@

\subsection{Matrix generators group $U_m$}

@@

\subsection{Owen Scrambling To The New Sequence}
